Figures and Plots

### 2025-07-05 docs/notebook/data_analysis.qmd
Updated Introduction/Methods completed Added all the codes and data files to the folder. Can publish directly from here for later..

```{r}
# install.packages("devtools")
#devtools::install_github("davidsjoberg/ggsankey")
library(ggsankey)

@@

```{r}
## For lnRR complete dataset
VCV_lnRR_published<- vcalc(vi = lnRR_variance,
                  cluster = paper_ID,
                  obs = Observation_ID,
                  data = dataset_lnRR_published,
                  rho = 0.5 # assuming that the effect sizes within the same study are correlated with rho = 0.5
                  )

all_in_lnRR <- rma.mv(yi = lnRR_sign, # specify lnRR as the effect size measure;
                    V = VCV_lnRR_published, # specify lnRR's sampling variance-covariance matrix;
                    mods = ~ - 1 + time_of_gnm_addition + 
                      year_MeanCenter + 
                      sqrt_effective_n_inv + 
                      comparision_type + 
                      bird_species + 
                      trait_type,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                    control = list(optimizer="optim"),
                    test = "t",
                    data = dataset_lnRR_published)

## For SMDH complete published dataset
VCV_SMDH_published<- vcalc(vi = SMDH_variance,
                  cluster = paper_ID,
                  obs = Observation_ID,
                  data = dataset_SMDH_published,
                  rho = 0.5 # assuming that the effect sizes within the same study are correlated with rho = 0.5
                  )

all_in_SMDH<- rma.mv(yi = SMDH_sign, # specify lnRR as the effect size measure;
                    V = VCV_SMDH_published, # specify lnRR's sampling variance-covariance matrix;
                    mods = ~ - 1 + time_of_gnm_addition + 
                      year_MeanCenter + 
                      sqrt_effective_n_inv + 
                      comparision_type + 
                      bird_species + 
                      trait_type,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                    control = list(optimizer="optim"),
                    test = "t",
                    data = dataset_SMDH_published)
```

>>>>

```{r}
# install.packages("devtools")
devtools::install_github("davidsjoberg/ggsankey")
library(ggsankey)

@@

```{r}
## For lnRR complete dataset

## I am changing the year of unpublished studies to 2025 which is to now, assuming that if they were to be published now, what would be the effect. I will keep in mind that in this model if I do find an evidence of publication bias over years, I should keep in mind it is non-independent of the unpublished studies

dataset_lnRR_all_in<-dataset_lnRR%>%mutate(year_publication = case_when(is.na(year_publication) ~ 2025, TRUE ~ year_publication))%>%
  mutate(year_MeanCenter=scale(as.numeric(year_publication),scale=F)[,1])%>%
  mutate(effective_n_inv=1/((4*effective_n_experiment*effective_n_control)/(effective_n_experiment+effective_n_control)))%>%
  mutate(sqrt_effective_n_inv=sqrt(effective_n_inv))%>%
  mutate(comparision_type= case_when(is.na(comparision_type) ~ 4, TRUE ~ comparision_type))

VCV_lnRR_all_in<- vcalc(vi = lnRR_variance,
                  cluster = paper_ID,
                  obs = Observation_ID,
                  data = dataset_lnRR_all_in,
                  rho = 0.5 # assuming that the effect sizes within the same study are correlated with rho = 0.5
                  )

lnRR_all_in <- rma.mv(yi = lnRR_sign, # specify lnRR as the effect size measure;
                    V = VCV_lnRR_all_in, # specify lnRR's sampling variance-covariance matrix;
                    mods = ~ time_of_gnm_addition + 
                      year_MeanCenter + 
                      sqrt_effective_n_inv + 
                      as.factor(comparision_type) + 
                      bird_species + 
                      trait_type,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                    control = list(optimizer="optim"),
                    test = "t",
                    data = dataset_lnRR_all_in)

## For SMDH complete published dataset

dataset_SMDH_all_in<-dataset_SMDH%>%mutate(year_publication = case_when(is.na(year_publication) ~ 2025, TRUE ~ year_publication))%>%
  mutate(year_MeanCenter=scale(as.numeric(year_publication),scale=F)[,1])%>%
  mutate(effective_n_inv=1/((4*effective_n_experiment*effective_n_control)/(effective_n_experiment+effective_n_control)))%>%
  mutate(sqrt_effective_n_inv=sqrt(effective_n_inv))%>%
  mutate(comparision_type= case_when(is.na(comparision_type) ~ 4, TRUE ~ comparision_type))

VCV_SMDH_all_in<- vcalc(vi = SMDH_variance,
                  cluster = paper_ID,
                  obs = Observation_ID,
                  data = dataset_SMDH_all_in,
                  rho = 0.5 # assuming that the effect sizes within the same study are correlated with rho = 0.5
                  )

SMDH_all_in<- rma.mv(yi = SMDH_sign, # specify lnRR as the effect size measure;
                    V = VCV_SMDH_all_in, # specify lnRR's sampling variance-covariance matrix;
                    mods = ~ time_of_gnm_addition + 
                      year_MeanCenter + 
                      sqrt_effective_n_inv + 
                      as.factor(comparision_type) + 
                      bird_species + 
                      trait_type,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                    control = list(optimizer="optim"),
                    test = "t",
                    data = dataset_SMDH_all_in)
```

### 2025-07-05 docs/notebook/figures.qmd
Updated Introduction/Methods completed Added all the codes and data files to the folder. Can publish directly from here for later..

```{r}
devtools::install_github("daniel1noble/orchaRd", ref = "main", force = TRUE)
pacman::p_load(ggplot2,dplyr,stringr,readr,cowplot)

@@

lnRR_design_plot <-
  orchard_plot(mod_results(
      lnRR_design,
      mod = "comparision_type",
      group = "paper_ID"
    ), xlab ="") +
  scale_colour_manual(values = c( "#218833","mediumpurple4","#895129")) +
  scale_fill_manual(values = c( "#218833","mediumpurple4","#895129")) +
  labs(x = "",
       y = "lnRR")+
    theme(
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.8),
    legend.position = "none",
    # legend.position.inside = c(0.98, 0.05),
    # legend.justification = c(1, 0), 
    # legend.direction = "horizontal",
    legend.title = element_text(size = 7),
    legend.text = element_text(size = 7),
    axis.title = element_text(size = 12),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(angle = 0, vjust = 0.5, hjust = 1)
  ) + scale_x_discrete(labels = c("Non-aromatic Vs. Aromatic","Aromatic Vs. Blank", "Non-Aromatic Vs. Blank"))

SMDH_design_plot <-
  orchard_plot(mod_results(
      SMDH_design,
      mod = "comparision_type",
      group = "paper_ID"
    ), xlab ="") +
  scale_colour_manual(values = c( "#218833","mediumpurple4","#895129")) +
  scale_fill_manual(values = c( "#218833","mediumpurple4","#895129")) +
  labs(x = "",
       y = "SMDH")+
    theme(
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.8),
    legend.position = "none",
    # legend.position.inside = c(0.98, 0.05),
    # legend.justification = c(1, 0), 
    # legend.direction = "horizontal",
    legend.title = element_text(size = 7),
    legend.text = element_text(size = 7),
    axis.title = element_text(size = 12),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(angle = 0, vjust = 0.5, hjust = 1)
  )

@@

## I am going to remove the y ticks from the SMDH plots to make it smaller..
SMDH_design_plot <- SMDH_design_plot +
  theme(axis.text.y  = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks.y = element_blank())

SMDH_time_gnm_addition_plot <- SMDH_time_gnm_addition_plot +
  theme(axis.text.y  = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks.y = element_blank())

# Combine using patchwork
Figure_design_time_gnm_addition <- ((lnRR_design_plot + SMDH_design_plot) /
                               (lnRR_time_gnm_addition_plot + SMDH_time_gnm_addition_plot)) +
  plot_annotation(tag_levels = "A")

ggsave(here::here("figures/Figure_design_time_gnm_addition.png"), plot = Figure_design_time_gnm_addition, height = 300, width = 400, units = "mm", dpi = 600)

>>>>

```{r}
devtools::install_github("daniel1noble/orchaRd", ref = "main", force = TRUE)
pacman::p_load(ggplot2,dplyr,stringr,readr,cowplot,maps)

@@

# lnRR_design_plot <-
#   orchard_plot(mod_results(
#       lnRR_design,
#       mod = "comparision_type",
#       group = "paper_ID"
#     ), xlab ="") +
#   scale_colour_manual(values = c( "#218833","mediumpurple4","#895129")) +
#   scale_fill_manual(values = c( "#218833","mediumpurple4","#895129")) +
#   labs(x = "",
#        y = "lnRR")+
#     theme(
#     panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.8),
#     legend.position = "none",
#     # legend.position.inside = c(0.98, 0.05),
#     # legend.justification = c(1, 0), 
#     # legend.direction = "horizontal",
#     legend.title = element_text(size = 7),
#     legend.text = element_text(size = 7),
#     axis.title = element_text(size = 12),
#     axis.text.x = element_text(size = 12),
#     axis.text.y = element_text(angle = 0, vjust = 0.5, hjust = 1)
#   ) + scale_x_discrete(labels = c("Non-aromatic Vs. Aromatic","Blank Vs. Aromatic", "Blank Vs. Non-Aromatic"))
# 
# SMDH_design_plot <-
#   orchard_plot(mod_results(
#       SMDH_design,
#       mod = "comparision_type",
#       group = "paper_ID"
#     ), xlab ="") +
#   scale_colour_manual(values = c( "#218833","mediumpurple4","#895129")) +
#   scale_fill_manual(values = c( "#218833","mediumpurple4","#895129")) +
#   labs(x = "",
#        y = "SMDH")+
#     theme(
#     panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.8),
#     legend.position = "none",
#     # legend.position.inside = c(0.98, 0.05),
#     # legend.justification = c(1, 0), 
#     # legend.direction = "horizontal",
#     legend.title = element_text(size = 7),
#     legend.text = element_text(size = 7),
#     axis.title = element_text(size = 12),
#     axis.text.x = element_text(size = 12),
#     axis.text.y = element_text(angle = 0, vjust = 0.5, hjust = 1)
#   )

@@

# ## I am going to remove the y ticks from the SMDH plots to make it smaller..
# SMDH_design_plot <- SMDH_design_plot +
#   theme(axis.text.y  = element_blank(),
#         axis.title.y = element_blank(),
#         axis.ticks.y = element_blank())
# 
# SMDH_time_gnm_addition_plot <- SMDH_time_gnm_addition_plot +
#   theme(axis.text.y  = element_blank(),
#         axis.title.y = element_blank(),
#         axis.ticks.y = element_blank())
# 
# # Combine using patchwork
# Figure_design_time_gnm_addition <- ((lnRR_design_plot + SMDH_design_plot) /
#                                (lnRR_time_gnm_addition_plot + SMDH_time_gnm_addition_plot)) +
#   plot_annotation(tag_levels = "A")
# 
# ggsave(here::here("figures/Figure_design_time_gnm_addition.png"), plot = Figure_design_time_gnm_addition, height = 300, width = 400, units = "mm", dpi = 600)
#        
```
